Extending the WebID Protocol with Access Delegation

### 2012-08-04 paper.tex

The simplest solution for any user agent $A$ wishing to act as the user $U$, is for it to create a public/private key pair,  and with the public key thus generated create a certificate $C_u$ with $U$'s WebID in the Subject Alternative Name position and add the public key to the user $U$'s profile $P_u$.
Having done this the agent $A$ can then connect using the certificate $C_u$ and its associated private key, to any service it wishes to, when it wants to work for/as the user $U$.
This has the advantage of requiring no change to the WebID protocol.
It does give the agent $A$ all powers of the user $U$ and so this type of privilege should be limited to agents in which are either extensions of $U$ or ones $U$ can or must fully trust anyway.

@@

What kind of agents does it make sense to give such rights to?
Or put another way: what agent would giving such rights to not constitute a reduction in security to the user?
Or: which agent does the user and the authorization server already rely on completely?

In addition to the well known User Agents found in browsers on many different devices, there is one agent  on which any service requiring WebID authentication relies on completely and this is the WebID profile's origin server~\cite{barth-a-2011--a}.
Any server requiring authentication of an agent using WebID must trust the origin server serving the WebID Profile.
That origin server can at any time change the public keys in the profiles sent back to any authorization agents and indeed change any information in the profile it wants to.
Any user profile can then be thought of as being a user according to the origin server ie a \url{user@origin.org:port} - where \url{origin.org} is the domain of the origin server, and \url{port} is the port the server is running on.

When the user owns the origin server and there is a one to one relation between the two, as is principally the case in a personal FreedomBox, then the origin server is again just an extension of the user, and in that case there need be no thought of a difference between the user and the origin server.
In that case the origin server can in fact just authenticate as the user whenever needed.

It should be noted that each device can still be distinguished by servers in case of a problem by logging the exact public key that was used to connect to a service. 
It should be possible then to track down which device's private key had been compromised in case of unusual authentication behavior .

Things get more interesting when an origin server agent $A$ is serving a number of different profiles $P_1$ \ldots $P_n$ each identifying respectively users $U_1$ \ldots $U_n$ and needs to act on behalf of these users.
This would be the situation for servers acting for companies, government agencies, educational instituion, charities, football club, ....
This brings up two issues that need to be taken into account: first an issue of data perspective and an efficiency issue.

\paragraph{Keeping views distinct}-
First whenever $A$ connects as $U_1$ to a remote resource $R$ it has to place the data received in a seperate graph from the one it stores public non authenticated data in, and to the one it stores a representation to the same resurce $R$ seen when connected as the different user $U_2$. 
This is because the resource may return different representations depending on who is connecting to it - one representation for close friends of the owner of $R$ perhaps, and one for more distant ones. 
This can be illustrated using N3 graphs, by specifying a relation between a view on a resource (using a yet to be settled on vocabulary), and a graph which is the log:semantics of that view (cf. Listing~\ref{listing:views}). 
It is very important that the agent receiving information from different users not merge the information that was destined to different users, or else information leakage will severely reduce the trust other agents have in that server $A$. 
This requires carefully keeping identified information that was aimed at different users seperate, perhaps by creating different graph stores: one for publically - non authenticated - information, and one graph store for each user the server needs to authenticate as.

@@

\paragraph{Efficiency}-Secondly, whenever the Origin Server needs to act as $U_1$ to a remote server $R$ it will need to open a new TLS connection to that server using the certificate $C_{u1}$ for that user.
This could require $N$ parallel connections to the same server $S$. 
Opening TLS connections is somewhat expensive: it requires opening up a new TCP connection and starting an expensive asymmetric key cryptography handshake. 
For small numbers of N this can be defended, but as N becomes large the inefficiency this could create between two
organisations may be a serious drawback.
%\todo{calculate how inefficient this is, and at what size of network this starts really become problematic}

@@

The Capulet FB Guard having verified that this request does indeed come from the origin server of Romeo's profile
\lstinline|<https://montague.net/romeo>| would then be able to accept that since the requesting server could in any case act as Romeo whenever it chose to do so, this protocol making exchange between the Montague and Capulet housholds more efficient without making it less secure, this be justification enough for it (the capulet freedom box) to  serve the representation that Romeo would have received had he connected to the server directly.  
We will argue further down that this is very similar to the way CORS deals with access control, namely by adding a header to a request to clarify what agent is making the request.

This \lstinline|On-Behalf-Of| header makes it possible for the \textit{secretary} to open only one TLS connection to a remote server and be 
able to specify for each request on behalf of whome it is making it. The Montague Freedom Box could make a request to the capulet Freedom Box \lstinline|On-Behalf-Of| Romeo, a few milliseconds later \lstinline|On-Behalf-Of| Lord Montague, and so on... 
Again as explained in the previous section is is the responsibility of the Montague Freedom Box to keep the returned information seperate, and act with secrecy whenever it is acting on behalf of one or the other of the users. 
As such it's role is not unlike the role of Friar Laurence in the play - dedication to people's private matters requires tact and secrecy.

@@

If we now identify with a WebID the secretary that can act on behalf of a user, we can generalise the protocol somewhat.
That is, when an agent - such as the Montague Freedom Box Laurence with WebID \lstinline|<http://montague.net/fb#laurence>| - authenticates to a remote server it can use its own WebID. 
This would allow the Agent to have its own public key, and so for a compromise of the client public key not to affect the server public key.

But how would the remote server know that it can trust that secretary to be acting on behalf of a particular user?
It can no longer just compare the TLS keys of the requesting agent to see if it comes from the same Origin Server.
We need to make this relation explicity by use of a special RDF relation provisionally called \lstinline|:secretary|%
\footnote{A object property with a domain and range as \lstinline|foaf:Agent| which we will provisionally place
in the \lstinline|cert:| namespace, though it may be more appropriately placed in the \lstinline|auth:| namespace.}

@@

The remote server can then verify that the identified agent is the secretary of the agent he wishes to act on behalf of (as sepecified in the \lstinline|On-Behalf-Of| header, by dereferencing that user's profile and verifying that the user specifies the \lstinline|:secretary| relation there, as it would if Romeo had the foaf profiles as shown in Listing \ref{listing:webid1}.

@@

However, at this point we postpone the answer to the question of how we should describe the access control rules in a users WebID profile.
There already exists first attempts from the Linked Data/Semantic Web and the Read Write Web communities such as the WebAccessControl vocabulary\footnote{\url{http://www.w3.org/wiki/WebAccessControl}}, $_{dg}$FOAF~\cite{schwagereit-f-2010-181-a} as well as different projects on policy creation and management~\cite{kagal-l-2005--a}.
In our future work we will try to close the gap between the policy and access control vocabularies on the one hand and WebID authentication on the other hand.
A rough abstract in which direction this should be elaborated is presented in \autoref{listing:views} and we hope that we can sensitize others to take up this challenge too.

>>>>

The simplest solution for any user agent $A$ wishing to act as the user $U$, is for it to create a public/private key pair,  and with the public key thus generated create a certificate $C_u$ with $U$'s WebID in the Subject Alternative Name position and add the public key to the user $U$'s profile $P_u$.
Having done this the agent $A$ can then connect using the certificate $C_u$ and its associated private key, to any service it wishes to, when it wants to work for/as the user $U$.
This has the advantage of requiring no change to the WebID protocol.
It does give the agent $A$ all powers of the user $U$ and so this type of privilege should be limited to agents which are either extensions of $U$ or ones $U$ can or must fully trust anyway.

@@

%In addition to the well known User Agents found in browsers on many different devices, there is one agent on which any service requiring WebID authentication trusts implicitly and this is the WebID profile's origin server.
Any server requiring authentication of an agent using WebID must implicitly trust the origin server~\cite{barth-a-2011--a} hosting the WebID profile, since the profile hosting server can change any information in the profile.
%Any user profile can then be thought of as being a user according to the origin server, i.e. \textit{user@origin.org:port} - where \textit{origin.org} is the domain of the origin server, and \textit{port} is the port the server is running on.

When the user owns the origin server and there is a one-to-one relation between the two, as it is the case in a personal FreedomBox\footnote{http://www.freedomboxfoundation.org/}, then the origin server is just an extension of the user, and in that case trusting the server is the same as trusting the user.
%In that case the origin server can in fact just authenticate as the user whenever needed.

It should be noted that each device can still be distinguished by servers in case of a problem by logging the exact public key that was used to connect to a service. 
It should be possible then to track down which device's private key had been compromised in case of unusual behaviour .

Things become more interesting when an origin server agent $A$ is serving a number of different profiles $P_1$ \ldots $P_n$ each identifying respectively users $U_1$ \ldots $U_n$ and needs to act on behalf of each of these users.
This would be the situation for company servers, government agencies, educational institutions, charities, football clubs, etc.
This brings up two issues that need to be taken into account, an issue of data perspective and an issue of efficiency.

\paragraph{Keeping views distinct}-
First whenever $A$ connects as $U_1$ to a remote resource $R$ it has to place the data received in a separate graph from the one it stores public non authenticated data in, and to the one it stores a representation to the same resurce $R$ seen when connected as the different user $U_2$. 
This is because the resource may return different representations depending on who is connecting to it - one representation for close friends of the owner of $R$ perhaps, and one for more distant ones. 
This can be illustrated using N3 graphs, by specifying a relation between a view on a resource (using a yet to be settled on vocabulary), and a graph which is the log:semantics of that view (cf. Listing~\ref{listing:views}). 
It is very important that the agent receiving information from different users not merge the information that was destined to different users, or else information leakage will severely reduce the trust other agents have in that server $A$. 
This requires carefully keeping identified information that was aimed at different users seperate, perhaps by creating different graph stores: one for publically - non authenticated - information, and one graph store for each user the server needs to authenticate as.

@@

\paragraph{Efficiency}-Secondly, whenever the Origin Server needs to act as $U_1$ to a remote server $R$ it will need to open a new TLS connection to that server using the certificate $C_{u1}$ for that user.
This could require $N$ parallel connections to the same server $S$. 
Opening TLS connections is somewhat expensive, as it requires expensive asymmetric key cryptographic computations. 
When N becomes large, this inefficiency will be perceived as a serious drawback between two organisations.
%\todo{calculate how inefficient this is, and at what size of network this starts really become problematic}

@@

The Capulet FreedomBox Guard having verified that this request does indeed come from the origin server of Romeo's profile
https://montague.net/romeo would then be able to accept that since the requesting server could in any case act as Romeo whenever it chose to do so, this protocol making exchange between the Montague and Capulet houses more efficient without making it less secure, this be justification enough for it (the Capulet FreedomBox) to  serve the representation that Romeo would have received had he connected to the server directly.  
In Section~\ref{sec:relatedwork} we will argue that this is very similar to the way CORS deals with access control, namely by adding a header to a request in order to clarify what agent is making the request.

This \lstinline|On-Behalf-Of| header makes it possible for the \textit{secretary} to open only one TLS connection to a remote server and be 
able to specify for each request on behalf of whom it is making it. The Montague FreedomBox could make a request to the Capulet Freedom Box \lstinline|On-Behalf-Of| Romeo, a few milliseconds later \lstinline|On-Behalf-Of| Lord Montague, and so on. 
Again as explained in the previous section is is the responsibility of the Montague FreedomBox to keep the returned information separate, and act with secrecy whenever it is acting on behalf of one or the other of the users. 
As such it's role is not unlike the role of Friar Laurence in the play - dedication to people's private matters requires tact and secrecy.

@@

If we now identify with a WebID the secretary that can act on behalf of a user, we can generalise the protocol somewhat.
That is, when an agent - such as the Montague FreedomBox Laurence with WebID \lstinline|<http://montague.net/fb#laurence>| - authenticates to a remote server it can use its own WebID. 
This would allow the Agent to have its own public key, and so for a compromise of the client public key not to affect the server public key.

But how would the remote server know that it can trust that secretary to be acting on behalf of a particular user?
It can no longer just compare the TLS keys of the requesting agent to see if it comes from the same Origin Server.
We need to make this relation explicit by use of a special RDF relation provisionally called \lstinline|:secretary|%
\footnote{A object property with a domain and range as \lstinline|foaf:Agent| which we will provisionally place
in the \lstinline|cert:| namespace, though it may be more appropriately placed in the \lstinline|auth:| namespace.}

@@

The remote server can then verify that the identified agent is the secretary of the agent he wishes to act on behalf of (as specified in the \lstinline|On-Behalf-Of| header, by dereferencing that user's profile and verifying that the user specifies the \lstinline|:secretary| relation there, as it would if Romeo had the FOAF profiles as shown in Listing \ref{listing:webid1}.

@@

However, at this point we postpone the answer to the question of how we should describe the access control rules in a users WebID profile.
There already exists first attempts from the Linked Data / Semantic Web and the Read Write Web communities such as the WebAccessControl vocabulary\footnote{\url{http://www.w3.org/wiki/WebAccessControl}}, $_{dg}$FOAF~\cite{schwagereit-f-2010-181-a} as well as different projects on policy creation and management~\cite{kagal-l-2005--a}.
In our future work we will try to close the gap between the policy and access control vocabularies on the one hand and WebID authentication on the other hand.
A rough abstract in which direction this should be elaborated is presented in \autoref{listing:views} and we hope that we can sensitize others to take up this challenge too.
